---
title: "Week7B"
author: "Sharad Bajpai"
date: "11/3/2019"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

Installing/loading the latest installr package:
```{r}
library(installr)
```

ENVIRONMENT SETUP

Step 1:
```{r}
setwd("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Week7")
```

Step 2:
```{r}
library(bitops)
library(RCurl)
library(ROAuth)
library(rjson)
library(ndjson)
library(streamR)
library(twitteR)
library(installr)
```

Step 3:
```{r}
download.file(url="http://curl.haxx.se/ca/cacert.pem",
              destfile="cacert.pem")
```

Step 4:
```{r}
outFile <- "tweets_sample.json"
```

Step 5:
```{r}
requestURL <- "https://api.twitter.com/oauth/request_token"
accessURL <- "https://api.twitter.com/oauth/access_token"
authURL <- "https://api.twitter.com/oauth/authorize"
consumerKey <- "cSZAJsDxhGUFHPpjvC0VqobPD"
consumerSecret <- "xpDhczS91UBdvZPOpZc6IjyRmFAHfWpjRYFgAPcoILioQjeXnh"
accessToken <- "1188575616251125760-GKPSvo9G2LTVaHmJChLvIHAQz4gVpH"
accessTokenSecret<- "3zpoJufC4Mxa7c19wNybj8IdJRl89K8oNezNO1ebZyPbQ"
```

Step 6:

my_oauth <- OAuthFactory$new(consumerKey=consumerKey, consumerSecret=consumerSecret,
requestURL=requestURL,
accessURL=accessURL,
authURL=authURL)

my_oauth$handshake(cainfo="cacert.pem")



$$Solution1$$
The solution was to update R installing/loading the latest 

install.packages("installr") 

library(installr)

updateR() 

updating R and re-install the packages that I used for the Twitter API (TwitteR, streamR)

Output: You don't need to update R because you are using the latest version.

Date/Time: 10/30, 10 am And 10/25, 3pm



$$Solution2$$

install.packages("devtools")

library(devtools)

Output : Error in function (type, msg, asError = TRUE): Unknown SSL protocol error in connection to api.twitter.com:443

Date/Time: 10/30,10am And 10/25, 3pm


$$Solution 3$$

library(rsconnect)

system("echo | openssl s_client -connect https://developer.twitter.com/en/apps:443 | sed -ne '/-BEGIN CERTIFICATE-/,/-END CERTIFICATE-/p'")

rsconnect::addConnectServer(url="https://server.com", certificate = "cacert.pem", 'test server')

Output: [1] 127

Output: Server 'test server' added successfully: https://server.com/__api__

Date/Time: 10/30, 10 am And 10/25, 5pm

$$Solution 4$$

Tried regenerating accessToken, AccessTokenSecret, ConsumerKey and ConsumerKeySecret. Tried running all the solution codes again. That didnt help either.

Output: Error in function (type, msg, asError = TRUE)  : 
  Unknown SSL protocol error in connection to api.twitter.com:443

Date/Time: 10/31, 3pm And 10/26, 7pm


$$Solution 5$$

my_oauth <- list( consumerKey=consumerKey,
consumerSecret=consumerSecret,
requestURL=requestURL,
accessURL=accessURL,
authURL=authURL)

my_oauth$handshake(cainfo="cacert.pem")

Output: Attempt to appy non-funciton

Date/Time: 10/31, 3pm And 10/27, 2pm


$$Solution 6$$

library(RCurl)

system("echo | openssl s_client -connect https://developer.twitter.com/en/apps:443 | sed -ne '/-BEGIN CERTIFICATE-/,/-END CERTIFICATE-/p'")

Output - [1] 127

Date/time : 10/31, 4 pm


$$Solution 7$$

.opts = list(ssl.verifypeer = TRUE)

Output : Error in function (type, msg, asError = TRUE): Unknown SSL protocol error in connection to api.twitter.com:443

Date/Time: 10/31, 6 pm 


$$Solution 8$$

my_oauth <- OAuthFactory$new( consumerKey=consumerKey,
                            consumerSecret=consumerSecret,
                              requestURL=requestURL,
                              accessURL=accessURL,
                              authURL=authURL)

my_oauth$handshake(cainfo=system.file("CurlSSL","cacert.pem", package = "RCurl"))

Output: Error in function (type, msg, asError = TRUE) : Unknown SSL protocol error in connection to api.twitter.com:443

Date/Time: 10/31, 7 pm And 11/2, 4pm

$$Solution 9$$

curl_setopt($ch, CURLOPT_SSLVERSION,3)

Output: Error in function (type, msg, asError = TRUE) : Unknown SSL protocol error in connection to api.twitter.com:443

Date/Time: 11/2, 6 pm 

Step 7

setup_twitter_oauth(consumerKey, consumerSecret, accessToken,
accessTokenSecret)

GETTING TWEET DATA

Step 1:

sampleStream( file=outFile, oauth=my_oauth, tweets=100 )

Step 2:

follow <- ""

track <- "Boston,RedSoxs"

location <- c(23.786699, 60.878590, 37.097000, 77.840813)

filterStream( file.name=outFile, follow=follow, track=track, locations=location, oauth=my_oauth, timeout=5)

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```