---
title: "Week10"
author: "Sharad Bajpai"
date: "11/17/2019"
output: pdf_document
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

1. Revisit the census schema you created for homework 6. After installing SQLite, implement the tables for your database design in SQLite and load the data into the correct tables using either SQL INSERT statements or CSV loads. Make sure the database design is normalized (at least 3NF) and has minimal redundancy. Make sure your SQLite tables have primary keys as well as foreign keys for relationships.
```{r}
library(RSQLite)
library(tidyverse)
library(readr)
FipsEducationsDA5020 <- read_csv("C:/Users/bajpa/Downloads/Google Chrome downloads/DA5020/FipsEducationsDA5020.csv")
View(FipsEducationsDA5020)

library(readr)
FipsUnemploymentDA5020 <- read_csv("C:/Users/bajpa/Downloads/Google Chrome downloads/DA5020/FipsUnemploymentDA5020.csv")
View(FipsUnemploymentDA5020)

edu <- FipsEducationsDA5020 %>%
  spread(key = percent_measure, value = percent ) %>%
  separate(county_state, into = c("state","county"))

unemp <- FipsUnemploymentDA5020
colnames(unemp) <- c("fipsnumber","year1","percent_unemployed")

fips<-unique(cbind.data.frame((edu$fips),(edu$state),(edu$county)))
colnames(fips) <- c("fipsnumber","state","county")
rural_urban_code <- unique(cbind.data.frame(as.character(edu$rural_urban_cont_code),as.character(edu$description)))
colnames(rural_urban_code) <- c("rurl_urbn_code","description")
rural_urban_code <- as_tibble(rural_urban_code)

database <- dbConnect(SQLite(), dbname="Fips.sqlite")
dbWriteTable(conn = database, name = 'education', value = edu, row.names=F, header=T, overwrite=T)
dbWriteTable(conn = database, name = 'fips', value = fips, row.names=F, header=T, overwrite=T)
dbWriteTable(conn = database, name = "rural_urban", value = rural_urban_code, row.names=F, header=T, overwrite=T)
dbWriteTable(conn = database, name = "unemployment", value = unemp, row.names=F, header=T, overwrite=T)
dbListTables(database)

edu1 <- unique(cbind.data.frame(as.character(edu$fips), as.character(edu$year),
                                      as.character(edu$rural_urban_cont_code), as.character(edu$percent_four_plus_years_college),
                                      as.character(edu$percent_has_some_college), as.character(edu$percent_hs_diploma), as.character(edu$`percent_less than_hs_diploma`)))
colnames(edu1) <- c("fips","year", "ru_code","percent_four_plus_years_college",
                          "percent_has_some_college","percent_hs_diploma","percent_less_than_hs_diploma")
dbWriteTable(conn = database, name = "education1", value = edu1, row.names = F, overwrite=T)
head(dbReadTable(database,"education1")) 
head(dbGetQuery(database,"select fips,year,state,county,ru_code,description,percent_four_plus_years_college,percent_has_some_college,percent_hs_diploma,percent_less_than_hs_diploma
                      from education1
                      INNER JOIN fips on education1.fips=fips.fipsnumber
                      INNER JOIN rural_urban on education1.ru_code=rural_urban.rurl_urbn_code"),10)
```

Write SQL expressions to answer the following queries:
Question 2:

2.0 In the year 1970, what is the population percent that did not earn a high school diploma for the Nantucket county in Massachusetts? What about the year 2015?
```{r}  
dbGetQuery(database, "select year,state,county,[percent_less than_hs_diploma]
           from education 
           where (year = 1970 and county = 'Nantucket')")

dbGetQuery(database, "select year,state,county,[percent_less than_hs_diploma]
           from education 
           where (year = 2015 and county = 'Nantucket')")
```
The population not attending a high school diploma for the Nantucket county in Massachusetts in 1970 is 33.7%

The population not attending a high school diploma for the Nantucket county in Massachusetts in 2015 is 5.2%

2.1 What is the average population percentage that did not earn a high school diploma for the counties in Alabama for the year 2015?
```{r}
head(dbGetQuery(database, "select year,state,county,AVG([percent_less than_hs_diploma])
           from education 
           WHERE (year = 2015 and state = 'AL')
           GROUP BY county")) 
```

2.2 What is the average percentage of college graduates for the counties in the state of Massachusetts for the year 2015?
```{r}
head(dbGetQuery(database,"select year,state,county,AVG(percent_four_plus_years_college)
           from education 
           where (year = 2015 and state = 'MA')
           GROUP BY county"))
```
2.3 Determine the average percentage of the population that did not earn a high school diploma for the counties in Alabama for each year within the dataset. The result should return the calendar year and the average percentage drop out rate for that year.
```{r}
dbGetQuery(database, "select year,AVG([percent_less than_hs_diploma]) as Avg_drop_out
           from education  
           where state = 'AL'
           GROUP BY year")
```

2.4 What is the most common rural_urban code for the U.S. counties?
```{r}
dbGetQuery(database,"select rural_urban_cont_code, count(*) as rur_urb_count 
           from education
           GROUP BY rural_urban_cont_code
           ORDER BY count(*) DESC ")
```
The most common rural_urban code for US counties is 6

2.5 Which counties have not been coded with a rural urban code? Return a result that contains two fields: County, State for the counties that has not been assigned a rural urban code. Do not return duplicate values in the result. Order the result alphabetically by state.
```{r}
head(dbGetQuery(database, "select state, county
           from education  
           where rural_urban_cont_code = 'NULL'
           GROUP BY state, county")) 
```

2.6 What is the minimal percentage of college graduates for the counties in the state of Mississippi for the year 2010?
There is no data available for the year 2010 in this dataset
```{r}
dbGetQuery(database, "select year,state,county,MIN(percent_four_plus_years_college) 
           from education  
           where (year = 2010 and state = 'MS')")
```
2.7 Which state contains the most number of counties that have not been provided a rural urban code?
```{r}
head(dbGetQuery(database, "select state, county, count(*) as Total
           from education  
           where rural_urban_cont_code = 'NULL'
           GROUP BY state, county"))
```
2.8 In the year 2015, which fip counties, U.S. states contain a higher percentage of unemployed citizens than the percentage of college graduates? List the county name and the state name. Order the result alphabetically by state.
```{r}
head(dbGetQuery(database, "select state,county,percent_four_plus_years_college, percent_unemployed 
           from education  
           inner join unemployment on unemployment.fipsnumber = education.fips
           where (year = 2015 and(percent_unemployed > percent_four_plus_years_college))
           Group by year ,state, county"))
```
2.9 Return the county, U.S. state and year that contains the highest percentage of college graduates in this dataset?
```{r}
dbGetQuery(database,"select year,county,state, MAX(percent_four_plus_years_college) as Max_College_graduates 
           from education")
```

3. Compare your SQL SELECT statements to your dplyr statements written to answer the same questions. Do you have a preference between the two methods? State your reasons for your preference.
SQLite 
```{r}
dbGetQuery(database, "select year,state,county,[percent_less than_hs_diploma]
           from education 
           where (year = 1970 and county = 'Nantucket')")
```
dplyr
```{r}
edu %>% 
  select("year","state", "county", "percent_less than_hs_diploma") %>% 
  filter(year==1970,county=="Nantucket")

```
As the above example displays, both the codes are almost the same. I still prefer dplyr over SQL SELECT Statements. There are few reasons to why I prefer dplyr over SQL SELECT.

1. To use SQL Select, one needs to create a database frame which is a tedious task as compared to read.csv for dplyr.

2. According to me, data preparation using sql select is little more complex as compared to dplyr.

3. Creating a new column is intuitive in dplyr but requires far less code. Whereas in Sql Select it requires couple of steps including primary key assingment, 'insert into', 'from' and 'select' functions.

4. I am more versed with dplyr package which is another reason for me to choose dplyr.

5. dplyr can also be used for analysis whereas sql is not.



4.Write a R function named get_state_county_education_data_dplyr(edf, state), it accepts
a data frame containing education data and a state’s abbreviation for arguments and produces a chart that shows the change in education across time for each county in that state. Use dplyr to extract the data. Write a few R statements that call the function
with different state values.
```{r}
get_state_county_education_data_dplyr <- function(edf, State) {
  
  df <-   edf %>%
    filter(state == State) %>%
    select(year,county,state,percent_four_plus_years_college, percent_has_some_college, 
           percent_hs_diploma,`percent_less than_hs_diploma`)
  
  ggplot(data = df)+
    geom_point(mapping = aes(x = percent_four_plus_years_college, y = county))+
    facet_wrap(~ year, nrow =  1)
  
} 
get_state_county_education_data_dplyr(edf =  edu, State = 'MI')  
get_state_county_education_data_dplyr(edf =  edu, State = 'AK') 
```

5. Write a R function named get_state_county_education_data_sql(edSQL, state), it accepts a SQL database connection containing education data and a state’s
abbreviation for arguments and produces a chart that shows the change in education across time for each county in that state. Use SQL SELECT to extract the data from the database. Write a few R statements that call the function with different state values.
```{r}
get_state_county_education_data_sql <- function(edf, State) {
  
  d <- "select year,county,percent_four_plus_years_college from %s where state = '%s' "
  d <- sprintf(d,edf,State)
  d <-  dbGetQuery(database,d)
  
  ggplot(data = d)+
    geom_point(mapping = aes(x = percent_four_plus_years_college, y = county))+
    facet_wrap(~year, nrow = 1)
  
} 
get_state_county_education_data_sql(edf =  'education', State = 'MI')  
get_state_county_education_data_sql(edf =  'education', State = 'RI')
```

6. Write a R function named get_state_county_unemployment_data_dplyr(udf, state), it
accepts a data frame containing unemployment data and state’s abbreviation and produces a chart that shows the change in unemployment across time for each county in that state. Use dplyr to extract the data. Write a few R statements that call the
function with different state values.
```{r}
fips_un <- inner_join(fips,unemp)
get_state_county_unemployment_data_dplyr <- function(udf, State) {
  
  df1 <-   fips_un %>%
    filter(state == State) %>%
    select(year1,state,county,percent_unemployed)
  
  ggplot(data = df1)+
    geom_point(mapping = aes(x = percent_unemployed, y = county))+
    facet_wrap(~year1, nrow =  2)
  
}
get_state_county_unemployment_data_dplyr(udf = fips_un, State = 'LA' )  
get_state_county_unemployment_data_dplyr(udf = fips_un, State = 'NM' )
```


7. Write a R function named get_state_county_unemployment_data_sql(udfSQL, state), it accepts a SQL database oject containing unemployment data and state’s abbreviation and produces a chart that shows the change in education across time for each county in that state. Use SQL SELECT to extract the data. Write a few R statements that call the function with different state values.
```{r}
get_state_county_unemployment_data_sql <- function(udf, State) {
  
df2 <- "select year1,state,county,percent_unemployed from %s a inner join fips f
  on a.fipsnumber = f.fipsnumber where state = '%s' "
  
  df2 <- sprintf(df2, udf, State)
  
  df2 <- dbGetQuery(database, df2)
  
  ggplot(data = df2)+
  geom_point(mapping = aes(x = percent_unemployed, y = county))+
    facet_wrap(~year1, nrow =  2)
}  
get_state_county_unemployment_data_sql(udf = 'unemployment', State = 'NV')
get_state_county_unemployment_data_sql(udf = 'unemployment', State = 'NH')
```
```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```