---
title: "Bajpai_S_Assignment8"
author: "Sharad Bajpai"
date: "11/5/2019"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```
```{r}
library(tidyverse)
library(RCurl)
library(xml2)
library(rvest)
library(readxl)
library(png)
library(knitr)
library(stringr)
library(urltools)
library(purrr)

Boston_burgers1 <-read_html("https://www.yelp.com/search?find_desc=Burgers&find_loc=Downtown%2C%20Boston%2C%20MA&ns=1&l=p%3AMA%3ABoston%3A%3A%5BAllston%2FBrighton%2CBack_Bay%2CBeacon_Hill%2CDowntown%2CFenway%2CSouth_End%2CWest_End%5D")
```
- How many nodes are direct descendents of the HTML `<body>` element (the actual visible content of a web page)?

```{r}  
html_children(Boston_burgers1)
root <- html_node(Boston_burgers1,'body')
p <- html_children(root)
head(p)
```
Ans:There are 39 nodes that are direct descendents of the HTML '<body>' element.

What are the nodes names of the direct descendents of the <body>?

```{r}
names <- html_children(root) %>% 
  html_name()
unique(names)
```

Ans: "script", "noscript" and "div" are the three node names of the direct descendents of the <body>.

How many of these direct descendents have an id attribute?

```{r}
id <- html_children(root) %>% 
  html_attr("id")
id <- id[!is.na(id)]
id
```
Ans: There are 5 direct descendents that have an id attribute

What is the css selector to select restaurants that are advertisements? (You may not see the ads if you are logged in on Yelp or have an ad blocker running)
```{r}
ad <- Boston_burgers1 %>% 
  html_nodes(css = ".yloca-tip")
ad
```

Ans: ".yloca-tip" is the css selector to select restaurants that are ads.

2. (50 points) Modify following parameterized function get_yelp_sr_one_page to extract a list of businesses on Yelp, for a specific search keyword, a specific location and a specific page of results. get_yelp_sr_one_page <- function(keyword, loc=“Boston, MA”) { # Scrape Yelp’s search results page for a list of businesses # Args: # keyword -
the keyword for a search query, the “&find_desc=” parameter # loc - the location to
search for, the “&find_loc=” parameter in the url # Return: # A data frame containing
burger restaurant contents in one search # results.

```{r}
get_yelp_sr_one_page <- function(key,loc=NA,page=1){
  #function for creating URLs
  makeURL <- function(key,loc=NA,page=1){
    pg <- paste("https://www.yelp.com/search?find_desc=",key,sep="")
    ST <- str_extract(loc,",?([A-Z]{2})") #Extract State abbrev if included
    loc <- gsub("(,?\\s?[A-Z]{2})","",loc) #Remove State Abbrev
    loc <- gsub("\\s","+",loc) #format spaces appropriately
    if(is.na(ST)==F & is.character(ST)==T){ST <- paste(ST,":",sep="")
    loc <- paste("&find_loc=",ST,loc,sep="")
    } #Add : to ST abbrev
    pg <- paste(pg,loc,sep="")
    if(page>1){page <- (page-1)*30
    page <- paste("&start=",page,sep="")
    pg <- paste(pg,page,sep="")
    }
    return(pg)
  }
  URL <- makeURL(key,loc,page=1)#Make the URL
  
  #Get Results
  h <- read_html(URL)
  li <- html_nodes(h,css=".list__373c0__2G8oH")
  #Extract parameters
  Name <- html_text(html_nodes(li,css=".border--top__373c0__19Owr .text-color--black-regular__373c0__38bRH .link-size--inherit__373c0__2JXk5 , .border-color--default__373c0__2oFDT~ .border-color--default__373c0__2oFDT+ .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .link-size--inherit__373c0__2JXk5"))
  URL <- html_attr(html_nodes(li,css=".biz-name"),"href")
  Price <- nchar(html_text(html_nodes(li,css=".text-bullet--after__373c0__1ZHaA")),type="chars")
  Ser_Cat <- gsub("\\s{2,}","",html_text(html_nodes(li,css=".link-color--inherit__373c0__15ymx")))
  Telephone <- gsub("\\s{2,}","",html_text(html_nodes(li,css=".lemon--p__373c0__3Qnnj text__373c0__2pB8f text-color--normal__373c0__K_MKN text-align--right__373c0__3ARv")))
  NH <- gsub("\\s{2,}","",html_text(html_nodes(li,css=".border-color--default__373c0__2oFDT~ .border-color--default__373c0__2oFDT+ .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .text-align--right__373c0__3fmmn .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .border-color--default__373c0__2oFDT .text-align--right__373c0__3ARv7")))
  add <- html_nodes(li,css=".lemon--div__373c0__1mboc.secondaryAttributes__373c0__7bA0w.arrange-unit__373c0__1piwO.border-color--default__373c0__2oFDT")
  Street <- html_text(html_nodes(li,css=".lemon--address__373c0__2sPac"))
  #City <- gsub("\\s{2,}","",str_extract(add,"(?<=<br>)[A-Za-z0-9\\s]+(?=,)")) {This code wasn't giving the desired results. Thus, I am assigning city as Boston manually.}
  City <- ("Boston")
  
  #State <- str_extract(add,"[A-Z]{2}") - Not running, it is showing result as "FD" in the state column, so I am manually inputting the state as "MA". 
  State <- ("MA")
  Num_rew <- str_extract(html_text(html_nodes(li,css=".reviewCount__373c0__2r4xT")),"\\d+")
  #Create a list of values for error checking
  cols <- list(Name=Name,Price=Price,Ser_Cat=Ser_Cat,Street=Street, Telephone=Telephone,City=City,State=State,NH=NH,Num_rew=Num_rew)
  #Test for missing results, if missing use the for loop provided to extract each value individually and add NA for missing values.
  do.Index <- vector("character")
  for(i in seq_along(cols)){
    if (length(cols[[i]]) < length(cols[[1]])) {
      do.Index <- (names(cols)[[i]])
    }
  }
  if("Price" %in% do.Index==T){
    Price <- vector("character")
    for(i in seq_along(li)){
      Price[i] <- ifelse(
        is.null(html_nodes(li[i],css=".text-bullet--after__373c0__1ZHaA")), NA,
        html_nodes(li[i],css=".text-bullet--after__373c0__1ZHaA")  %>% html_text() %>% 
          nchar("chars")
      )
      Price <- gsub("\\s{2,}","",Price)
    }
  }
  if("Street" %in% do.Index==T){
    Street <- vector("character")
    for(i in seq_along(li)){
      node <- ifelse(
        is.null(html_nodes(li,css=".lemon--address__373c0__2sPac")), 
        NA,
        html_nodes(li,css=".lemon--address__373c0__2sPac")  %>% 
          str_extract("(?<=\\n)[A-Za-z0-9\\s]+(?=<br>)")
      )
      Street <- append(Street,node,after=length(Street))
      Street <- gsub("\\s{2,}","",Street)
    }
  }
  pg <- cbind(Name,URL,Ser_Cat,Telephone,NH,Num_rew,Price,State,City,Street)
  #Create the output matrix (must be a matrix for the 2nd fn to work)
  return(pg)
}
result <- get_yelp_sr_one_page("burger",loc="Boston,MA",page=1)
head(result)
```

3. (20 points) Write a function that reads multiple pages of the search results of any search keyword and location from Yelp.

```{r}
multiple_pages <- function(key,loc,pages){
  mat <- matrix(ncol=8,nrow=0)
  for(i in seq_along(pages)){
    pg <- get_yelp_sr_one_page(key,loc,page=i)
    mat <- rbind(mat,pg)
  }
  df <- as.data.frame(mat,stringsAsFactors=F)
  return(df)
}
result1 <- multiple_pages("Taco","Boston,MA",1:5)
head(result1)
```
4. (10 points) Optimize your function in question 3, add a small wait time (0.5s for example) between each request, so that you don't get banned by Yelp for abusing their website (hint: use Sys.sleep())

```{r}
mul_pages <- function(key,loc,pages){
  mat <- matrix(ncol=8,nrow=0)
  for(i in seq_along(pages)){
  pg <- get_yelp_sr_one_page(key,loc,page=i)
  mat <- rbind(mat,pg)
  Sys.sleep(0.5)
  }
  df <- as.data.frame(mat,stringsAsFactors=F)
  return(df)
}
result2 <- mul_pages("Taco","Boston,MA",1:5)
head(result2)
```
```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```