---
title: "Term Project"
author: "Sharad Bajpai"
date: "11/23/2019"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

Libraries required

```{r}
library(tidyverse)
library(rvest)
library(knitr)
library(readxl)
library(png)
```
Importing datasets for Private Insurance:
```{r}
library(readr)
California <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Private/California.csv")
head(California)

library(readr)
NewYork <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Private/NewYork.csv")
head(NewYork)

library(readr)
Pennsylvania <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Private/Pennsylvania.csv")
head(Pennsylvania)

library(readr)
Virginia <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Private/Virginia.csv")
head(Virginia)

library(readr)
Wisconsin <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Private/Wisconsin.csv")
head(Wisconsin)
```
Cleaning the State column name:
```{r}
California$State<-gsub("[[:upper:]]","",California$State)
California$State<-'CA'
head(California)

NewYork$State<-gsub("[[:upper:]]","",NewYork$State)
NewYork$State<-'NY'
head(NewYork)

Pennsylvania$State<-gsub("[[:upper:]]","",Pennsylvania$State)
Pennsylvania$State<-'PA'
head(Pennsylvania)

Virginia$State<-gsub("[[:upper:]]","",Virginia$State)
Virginia$State<-'VA'
head(Virginia)

Wisconsin$State<-gsub("[[:upper:]]","",Wisconsin$State)
Wisconsin$State<-'WI'
head(Wisconsin)
```

Combining these datasets into one.
```{r}
Insurance<-rbind(California,NewYork,Pennsylvania,Virginia, Wisconsin)
head(Insurance)
Ins_Id <-1
Insurance <- cbind(Ins_Id,Insurance)
HealthPlanId<- c(101:200)
HealthPlanId<-as.data.frame(HealthPlanId)
Insurance <- cbind(HealthPlanId,Insurance)
head(Insurance, 8)
```

Removing Special Characters from the columns:
```{r}
Insurance$`Consumer Satisfaction`<-gsub("I","0",Insurance$`Consumer Satisfaction`)
Insurance$Prevention<-gsub("I","0", Insurance$Prevention)
Insurance$Treatment<-gsub("I","0", Insurance$Treatment)
Insurance$`Overall Rating`<-as.numeric(Insurance$`Overall Rating`)
Insurance$`Consumer Satisfaction`<-as.numeric(Insurance$`Consumer Satisfaction`)
Insurance$Prevention<-as.numeric(Insurance$Prevention)
Insurance$Treatment<-as.numeric(Insurance$Treatment)
head(Insurance)
```

Importing dataset for Medicare:
```{r}
library(readr)
CaliforniaMedicare <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Medicare/CaliforniaMedicare.csv")
head(CaliforniaMedicare)

library(readr)
NewYorkMedicare <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Medicare/NewYorkMedicare.csv")
head(NewYorkMedicare)

library(readr)
PennsylvaniaMedicare <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Medicare/PennsylvaniaMedicare.csv")
head(PennsylvaniaMedicare)

library(readr)
VirginiaMedicare <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Medicare/VirginiaMedicare.csv")
head(VirginiaMedicare)

library(readr)
WisconsinMedicare <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Medicare/WisconsinMedicare.csv")
head(WisconsinMedicare)
```
Cleaning the State column name:
```{r}
CaliforniaMedicare$State<-gsub("[[:upper:]]","",CaliforniaMedicare$State)
CaliforniaMedicare$State<-'CA'
head(California)

NewYorkMedicare$State<-gsub("[[:upper:]]","",NewYorkMedicare$State)
NewYorkMedicare$State<-'NY'
head(NewYorkMedicare)

PennsylvaniaMedicare$State<-gsub("[[:upper:]]","",PennsylvaniaMedicare$State)
PennsylvaniaMedicare$State<-'PA'
head(PennsylvaniaMedicare)

VirginiaMedicare$State<-gsub("[[:upper:]]","",VirginiaMedicare$State)
VirginiaMedicare$State<-'VA'
head(VirginiaMedicare)

WisconsinMedicare$State<-gsub("[[:upper:]]","",WisconsinMedicare$State)
WisconsinMedicare$State<-'WI'
head(WisconsinMedicare)
```
Combining these Medicare datasets into one.
```{r}
Medicare<-rbind(CaliforniaMedicare,NewYorkMedicare,PennsylvaniaMedicare,VirginiaMedicare, WisconsinMedicare)
head(Medicare)
Ins_Id<-2
Medicare<-cbind(Ins_Id,Medicare)
HealthPlanId<- c(1001:1099)
HealthPlanId<-as.data.frame(HealthPlanId)
Medicare <- cbind(HealthPlanId,Medicare)
head(Medicare)
```
Removing Special Characters from the columns:
```{r}
Medicare$`Consumer Satisfaction`<-gsub("I","0",Medicare$`Consumer Satisfaction`)
Medicare$Prevention<-gsub("I|0","0", Medicare$Prevention)
Medicare$Treatment<-gsub("I","0", Medicare$Treatment)
Medicare$`Overall Rating` <- gsub("Partial Data Reported|No Data Reported","0", Medicare$`Overall Rating`)
Medicare$`Overall Rating`<-as.numeric(Medicare$`Overall Rating`)
Medicare$`Consumer Satisfaction`<-as.numeric(Medicare$`Consumer Satisfaction`)
Medicare$Prevention<-as.numeric(Medicare$Prevention)
Medicare$Treatment<-as.numeric(Medicare$Treatment)
head(Medicare)
```

Fixing multiple values in Consumer Satsifaction column:
```{r}
Medicare$`Consumer Satisfaction`[17]<-1.5
Medicare$`Consumer Satisfaction`[20]<-2.0
head(Medicare)
```

Importing dataset for Medicaid:
```{r}
library(readr)
CaliforniaMedicaid <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Medicaid/CaliforniaMedicaid.csv")
head(CaliforniaMedicaid)

library(readr)
NewYorkMedicaid <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Medicaid/NewYorkMedicaid.csv")
head(NewYorkMedicaid)

library(readr)
PennsylvaniaMedicaid <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Medicaid/PennsylvaniaMedicaid.csv")
head(PennsylvaniaMedicaid)

library(readr)
VirginiaMedicaid <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Medicaid/VirginiaMedicaid.csv")
head(VirginiaMedicaid)

library(readr)
WisconsinMedicaid <- read_csv("C:/Users/bajpa/Downloads/Collecting, Storing, Retrieving Data/Project/Data/Medicaid/WisconsinMedicaid.csv")
head(WisconsinMedicaid)
```

Cleaning the State column name:
```{r}
CaliforniaMedicaid$State<-gsub("[[:upper:]]","",CaliforniaMedicaid$State)
CaliforniaMedicaid$State<-'CA'
head(California)

NewYorkMedicaid$State<-gsub("[[:upper:]]","",NewYorkMedicaid$State)
NewYorkMedicaid$State<-'NY'
head(NewYorkMedicaid)

PennsylvaniaMedicaid$State<-gsub("[[:upper:]]","",PennsylvaniaMedicaid$State)
PennsylvaniaMedicaid$State<-'PA'
head(PennsylvaniaMedicaid)

VirginiaMedicaid$State<-gsub("[[:upper:]]","",VirginiaMedicaid$State)
VirginiaMedicaid$State<-'VA'
head(VirginiaMedicaid)

WisconsinMedicaid$State<-gsub("[[:upper:]]","",WisconsinMedicaid$State)
WisconsinMedicaid$State<-'WI'
head(WisconsinMedicaid)
```
Combining these datasets into one.
```{r}
Medicaid<-rbind(CaliforniaMedicaid,NewYorkMedicaid,PennsylvaniaMedicaid,VirginiaMedicaid, WisconsinMedicaid)
head(Medicaid)
Ins_Id<-3
Medicaid<-cbind(Ins_Id,Medicaid)
HealthPlanId<- c(5001:5073)
HealthPlanId<-as.data.frame(HealthPlanId)
Medicaid <- cbind(HealthPlanId,Medicaid)
head(Medicaid)
```
Removing Special Characters from the columns:
```{r}
Medicaid$`Consumer Satisfaction`<-gsub("I","0",Medicaid$`Consumer Satisfaction`)
Medicaid$Prevention<-gsub("I|0","0", Medicaid$Prevention)
Medicaid$Treatment<-gsub("I","0", Medicaid$Treatment)
Medicaid$`Overall Rating` <- gsub("Partial Data Reported|No Data Reported","0", Medicaid$`Overall Rating`)
Medicaid$NCQA<-gsub("(In Process)|(Scheduled)|(Interim)|(In process)|[[:punct:]]"," ", Medicaid$NCQA)
Medicaid$`Overall Rating`<-as.numeric(Medicaid$`Overall Rating`)
Medicaid$`Consumer Satisfaction`<-as.numeric(Medicaid$`Consumer Satisfaction`)
Medicaid$Prevention<-as.numeric(Medicaid$Prevention)
Medicaid$Treatment<-as.numeric(Medicaid$Treatment)
head(Medicaid)
```

Fixing multiple values in Prevention column:
```{r}
Medicaid$Prevention[67]<-0.5
view(Medicaid)
```
Creating the Central Entity that connects all the other tables.
```{r}
Ins_Id<-c(1,2,3)
Ins_type<-c("Private_Insurance", "Medicare", "Medicaid")
Ins <- cbind(Ins_Id,Ins_type)
Ins<-as.data.frame(Ins)
head(Ins)
```

The four dataframes are Insurnace (Private Insurnace), Medicare and Medicaid and Ins
```{r}
view(Insurance)
view(Medicare)
view(Medicaid)
view(Ins)
```

Connecting to database:
```{r}
library(RSQLite)
database <- dbConnect(SQLite(), dbname="US_Insurance")
dbWriteTable(conn = database, name = 'Insurance', value = Insurance, row.names=F, header=T, overwrite=T)
dbWriteTable(conn = database, name = 'Medicare', value = Medicare, row.names=F, header=T, overwrite=T)
dbWriteTable(conn = database, name = "Medicaid", value = Medicaid, row.names=F, header=T, overwrite=T)
dbWriteTable(conn = database, name = "Ins", value = Ins, row.names=F, header=T, overwrite=T)
dbListTables(database)
```
Checking the database by running custom queries:
```{r}
dbGetQuery(database, "select Max([Overall Rating]), [Plan Name], State
           from Insurance 
           where (State='CA')")
```


```{r}
dbGetQuery(database, "select [Overall Rating], [Plan Name], State, Type, NCQA, [Consumer Satisfaction], [Prevention], Treatment
           from Insurance 
           where ([Consumer Satisfaction]>3.5 AND [Consumer Satisfaction]!='NA')")
```
```{r}
dbGetQuery(database, "select [Overall Rating], [Plan Name], State, Type, NCQA, [Consumer Satisfaction], [Prevention], Treatment
           from Medicare 
           where ([Overall Rating]>4 AND [Consumer Satisfaction]>4 AND[Consumer Satisfaction]!='NA') AND [Overall Rating]!='NA' ")
```
```{r}
dbGetQuery(database, "select [Overall Rating], [Plan Name], State, Type, NCQA, [Consumer Satisfaction], [Prevention], Treatment
           from Medicare 
           where ([Treatment]>4 AND [Prevention]>4 AND [Treatment]!='NA')")
```
```{r}
dbGetQuery(database, "select [Overall Rating], [Plan Name], State, Type, NCQA, [Consumer Satisfaction], [Prevention], Treatment
           from Medicaid
           where ([Overall Rating]>4 AND [Prevention]<4 AND [Prevention]!='NA' AND [Overall Rating]!='NA')")
```
```{r}
dbGetQuery(database, "select [Overall Rating], [Plan Name], State, Type, NCQA, [Consumer Satisfaction], [Prevention], Treatment
           from Medicaid
           where ([Overall Rating]=4.5 AND [Overall Rating]!='NA') AND State='NY'")
```

```{r}
dbGetQuery(database, "select *
                      from Ins
           ")
```
As you can see, I have tested all the four tables and the database is working efficiently.
Now I wish to carry out analysis to answer my research questions,
1. Average State ratings (of all the three insurance types) based on consumer satisfaction, treatment, and prevention.
Private Insurance Analysis
```{r}
B<- Insurance
B<-na.omit(B)
view(B)

b <- B %>% 
  na.omit() %>% 
  group_by(State) %>%
  summarise_at(vars(-NCQA, -'Plan Name',-Type,-HealthPlanId, -Ins_Id),funs(mean(., na.rm=TRUE)))
view(b)

Ans1_a<- b[order(-b$`Overall Rating`),]
head(Ans1_a)
```
In the Private Insurance analysis, Pennsylvania has the highest overall, consumer satisfaction and prevention rating whereas, Wisconsin has the highest treatment rating.

Medicare Analysis:
```{r}
C<- Medicare
C<-na.omit(C)
view(C)

c <- C %>% 
  na.omit() %>% 
  group_by(State) %>%
  summarise_at(vars(-NCQA, -'Plan Name',-Type, -HealthPlanId, -Ins_Id),funs(mean(., na.rm=TRUE)))
view(c)

Ans1_b<- c[order(-c$`Overall Rating`),]
head(Ans1_b)
```
In the Medicare analysis, New York has the highest overall and prevention rating whereas, Wisconsin has the highest treatment rating and Pennsylvania has highest consumer satisfaction rating.

Medicaid Analysis:
```{r}
D<- Medicaid
D<-na.omit(D)
view(D)

d <- D %>% 
  na.omit() %>% 
  group_by(State) %>%
  summarise_at(vars(-NCQA, -'Plan Name',-Type,-HealthPlanId, -Ins_Id),funs(mean(., na.rm=TRUE)))
view(d)

Ans1_c<- d[order(-d$`Overall Rating`),]
head(Ans1_c)
```
In the Medicaid Analysis, Pennsylvania has the highest Overall Rating and Treatment rating, Virginia has the highest Consumer satisfaction ratings whereas New York has the highest prevention rating.

2. Which insurance plan provides the best customer satisfaction, treatment, and prevention?
Health Plans for the Private Insurance:
```{r}
Ans2_a_1 <- B[order(-B$`Consumer Satisfaction`),]
Ans2_a_1 <- Ans2_a_1 %>% 
  select(`Plan Name`, State, `Consumer Satisfaction`)
head(Ans2_a_1,4)

Ans2_a_2 <- B[order(-B$Prevention),]
Ans2_a_2 <- Ans2_a_2 %>% 
  select(`Plan Name`, State, Prevention)
head(Ans2_a_2,4)

Ans2_a_3 <- B[order(-B$Treatment),]
Ans2_a_3 <- Ans2_a_3 %>% 
  select(`Plan Name`, State, Treatment)
head(Ans2_a_3,4)
```
Health Plan for Medicare:
```{r}

Ans2_b_1 <- C[order(-C$`Consumer Satisfaction`),]
Ans2_b_1 <- Ans2_b_1 %>% 
  select(`Plan Name`, State, `Consumer Satisfaction`)
head(Ans2_b_1,4)

Ans2_b_2 <- C[order(-C$Prevention),]
Ans2_b_2 <- Ans2_b_2 %>% 
  select(`Plan Name`, State, Prevention)
head(Ans2_b_2,4)

Ans2_b_3 <- C[order(-C$Treatment),]
Ans2_b_3 <- Ans2_b_3 %>% 
  select(`Plan Name`, State, Treatment)
head(Ans2_b_3,4)
```
Health Plan for Medicaid:
```{r}
Ans2_c_1 <- D[order(-D$`Consumer Satisfaction`),]
Ans2_c_1 <- Ans2_c_1 %>% 
  select(`Plan Name`, State, `Consumer Satisfaction`)
head(Ans2_c_1,4)

Ans2_c_2 <- D[order(-D$Prevention),]
Ans2_c_2 <- Ans2_c_2 %>% 
  select(`Plan Name`, State, Prevention)
head(Ans2_c_2,4)

Ans2_c_3 <- D[order(-D$Treatment),]
Ans2_c_3 <- Ans2_c_3 %>% 
  select(`Plan Name`, State, Treatment)
head(Ans2_c_3,4)
```
```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```